---
title: "Kenya Household Assets"
author: "Mburu"
date: "7/7/2020"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: no
    theme: united
    highlight: pygments
    css: github.css
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Kenya selected households assets 2019 census

I figured that it will be important for me to do this to show why it is impossible for online learning to be adopted in Kenya.

I used 2019 census data set which can be found [here](https://open.africa/dataset/2019-kenya-population-and-housing-census/resource/95eb7a07-71dd-48be-8c82-dd321ab329f1) and the counties shape file can be found [here](https://africaopendata.org/dataset/kenya-counties-shapefile)


```{r cars}
#Packages used

library(plotly)
library(tidyverse)
library(data.table)
library(sf)
library(DT)
library(tmap)
library(ggthemes)
```

# Read data set

```{r} 
household_assets <- fread("percentage-distribution-of-conventional-households-by-ownership-of-selected-household-assets-201.csv")

#
kenya_shapefile <- st_read("County") %>% setDT()

kenya_shapefile[, county := tolower(COUNTY)]

# rename column
setnames(household_assets, 
         c("County / Sub-County", "Conventional Households"  ), 
         c("county", "households"))
```

# Some minor cleaning


```{r}

# remove commas 
household_assets[, households := as.numeric(gsub(",|AR K    ", "", households))]

household_assets[,  county := tolower(county)]

sub_county <- household_assets %>% 
    group_by(county) %>%
    filter(households == max(households)) %>% setDT()


sub_county_melt <- melt(sub_county, 
                        id.vars = c("county", "households"))


```

# % of households with various assets 2019 census

- This is overall data set for the whole country computer devices is ownership about 8.8% this just means that about 91% of the students can't access online learning. This is is just a naive estimation the number could be higher.


```{r, fig.width=10, fig.height=4.5}
kenya_dat <- sub_county_melt[county == "kenya"]

p <- ggplot(kenya_dat, aes(variable, value)) +
    geom_bar(stat = "identity", width = 0.5, fill ="slategray2"  ) +
    geom_text(aes(variable, value, label = paste0(value, "%")),
              position = position_dodge(width = 0.5),
              vjust = 0.001, size = 4)+
    labs(x = "Household assets", y = "%",
         title = "% of households with various assets 2019 census",
         caption = "twitter\n@mmburu_w")+
    theme_fivethirtyeight()
p
```

# Merge shapefile with asset data sets

```{r}
 
sub_county_melt[county == "elgeyo/marakwet", county := "keiyo-marakwet"]
sub_county_melt[county == "tharaka-nithi", county := "tharaka"]
sub_county_melt[county ==  "taita/taveta", county := "taita taveta"]
sub_county_melt[county ==  "nairobi city", county := "nairobi"]

county_shapes <- merge(kenya_shapefile, sub_county_melt, by = "county") 

setnames(county_shapes, "value", "Percentage")
```

# Computer devices data

```{r}

computer <- county_shapes[variable  %in% c("Desk Top\nComputer/\nLaptop/ Tablet")]

# this converts to sf object
computer <- st_set_geometry(computer, "geometry")
```


# Percentage of households with computer devices per county

- That is if a household owns a tablet, laptop or a desktop


```{r, fig.width=7, fig.height=6}
#ttm()
tm_shape(computer)+
    tm_borders(col = "grey")+
    tm_fill(col = "Percentage")+
    tm_layout(title = "% of households with laptop/tablet/desk top \n 2019 census",
              title.size = 1, title.position = c(0.3, 0.95))
#ttm()
```


# % of households that can access internet

- This looks like is internet access through mobile phones

```{r, fig.width=7, fig.height=6}
internet <- county_shapes[variable == "Internet"]

internet <- st_set_geometry(internet, "geometry")

#ttm()
tm_shape(internet)+
    tm_borders(col = "grey")+
    tm_fill(col = "Percentage")+
    tm_layout(title = "% of households with internet acess",
              title.size = 1, title.position = c(0.3, 0.95))
```
