---
title: "Association Analysis: Finding Patterns in Baskets"
date: "2018-11-28"
description: "A teaching note on support, confidence, and lift using bakery transaction data to explain how association rules turn baskets into evidence."
permalink: /posts/data-evidence/association-analysis/association-analysis.html
share-permalink: "https://m-mburu.github.io/posts/data-evidence/association-analysis/association-analysis.html"
categories: ["Data & Evidence", "Teaching"]
format: html
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

data_table <- function(df){
    library(DT)
    n_col = ncol(df)-1
    datatable(df,
              rownames = FALSE,
              style = "bootstrap4", class = 'cell-border stripe',
              options = list(scrollX = TRUE,
                             columnDefs = list(list(className = 'dt-center', 
                                                    targets = 0:n_col))
              )
    )
}


```

# Association Analysis

Maybe you have heard of a grocery store that carried out analysis and found out that men who buy diapers between 5pm to 7pm were more likely to buy beer. The grocery store then moved the beer isle closer to the diaper isle and beer sales increased by 35%. This is called association analysis which was motivated by retail store data.

In this blog I will explore the basics of association analysis. The goal is to find out:


- **Support** measures how often items appear together. Suppose there are 10 transactions and 3 of them contain maize flour, rice, and bread. Then the support of that itemset is $3/10 = 0.3$. This is just the marginal probability that these items were bought together, or equivalently the share of transactions containing them.

  For the rule $\{\text{bread}, \text{maize flour}\} \rightarrow \{\text{rice}\}$, support can be written as:

  $$
  \operatorname{Support}(\{\text{bread}, \text{maize flour}\} \rightarrow \{\text{rice}\})
  =
  P(\text{bread}, \text{maize flour}, \text{rice}).
  $$

  More generally, if $X$ and $Y$ are itemsets, then:

  $$
  \operatorname{Support}(X \rightarrow Y) = P(X \cup Y).
  $$

- **Confidence** measures how likely one set of items is to appear once another set has already appeared. Suppose the proportion of transactions containing maize flour and bread is $0.4$, while the proportion containing maize flour, bread, and rice is $0.3$. Then the confidence is:

  $$
  \operatorname{Confidence}(\{\text{bread}, \text{maize flour}\} \rightarrow \{\text{rice}\})
  =
  \frac{\operatorname{Support}(\{\text{bread}, \text{maize flour}, \text{rice}\})}{\operatorname{Support}(\{\text{bread}, \text{maize flour}\})}
  =
  \frac{0.3}{0.4}
  =
  0.75.
  $$

  In other words, 75% of the transactions that contained bread and maize flour also contained rice. In general,

  $$
  \operatorname{Confidence}(X \rightarrow Y)
  =
  \frac{\operatorname{Support}(X \cup Y)}{\operatorname{Support}(X)}
  =
  P(Y \mid X).
  $$

- **Lift** measures how much more likely $Y$ is to be purchased once $X$ has been purchased, compared with the overall chance of purchasing $Y$.

  $$
  \operatorname{Lift}(X \rightarrow Y)
  =
  \frac{\operatorname{Confidence}(X \rightarrow Y)}{\operatorname{Support}(Y)}
  =
  \frac{P(Y \mid X)}{P(Y)}.
  $$

  A lift greater than $1$ suggests that $X$ and $Y$ occur together more often than we would expect by chance alone.

To make sense of these ideas, I will use a bakery dataset to work through association rules manually and then compare the results with the R package `arules`, which implements the Apriori algorithm for finding association rules. The dataset is available on Kaggle as BreadBasket_DMS. We start by first having a glimpse of this dataset.

```{r}

library(tidyverse)
library(knitr)
library(data.table)
library(DT)

dat <- setDT(read.csv( "BreadBasket_DMS.csv" ))

dat <- dat[Item != "NONE"]
head(dat[sample(1:nrow(dat), 10)]) %>% data_table()


```

## Data Cleaning and Exploratory Data Analysis

First step is to transform the data set into wide format. Column headers will be items sold in the bakery and the rows will be populated with 1 and 0 indicating whether that item was bought for that transaction.

```{r}

dat2 <- dcast(Date+Time+Transaction~Item, data = dat, fun.aggregate = length)
#dat2[, NONE := NULL]

sample_cols <- sample(4:ncol(dat2), 5)

item_names <- names(dat2)[4:97]

dat2[, (item_names) := lapply(.SD, function(x) ifelse(x==0, 0, 1)), .SDcols = item_names]

head(dat2[, c(1:3, sample_cols), with = F]) %>% data_table()
```

## How many items each customer buy?

On average a transaction has 2 items. The median is also 2, this shows that atleast 50% of the transactions contained 2 or more items and atleast 25% of the transactions have 1 item.

```{r}

number_items <- rowSums(dat2[, 4:97, with =F])

dat2[, number_items := number_items]

hist(number_items, col = "black", main = "Number of items bought")

sumStats <- dat2 %>%
    summarise(Average = round(mean(number_items), 2), Stdev = round(sd(number_items), 2),
              Median= median(number_items),
              Minimum = min(number_items), Maximum = max(number_items),
              First_qurtile = quantile(number_items,0.25,  na.rm = T),
              Third_qurtile=quantile(number_items,0.75,  na.rm = T))

data_table(sumStats)
```

## Top 10 bought items

Table below shows top ten most bought items and about 47.84% of the transactions contained coffee. Coffee was the most popular item in this bakery followed by bread.

```{r}



n_transacts_item_in <- colSums(dat2[, item_names, with = F])

data.frame(item = names(n_transacts_item_in),
           number =n_transacts_item_in) %>%
    mutate(Percentage = round(number/nrow(dat2)*100, 2)) %>%
    arrange(desc(number)) %>% head(10) %>% data_table()


```

Since we have transformed the data in the wide format and every transaction is in it's row we can visualize how the baskets look like. This is done by extracting the column names for the transactions where the value is 1. For each transaction, 1 represent that item being in that transaction.

```{r}


items_bought <- apply(dat2[, 4:97, with =F], 1, paste, collapse = "", sep = "")

list_items <- vector(mode = "list", length = length(items_bought))
for (i in 1:length(items_bought)) {
    index <- unlist(gregexpr("1", items_bought[i]))
    items_transaction_i <- item_names[index]
    items_transaction_i <- paste(items_transaction_i, sep = " ", collapse = " , ")
    list_items[[i]] <- items_transaction_i
}

head(unlist(list_items))
```

Data frame below shows how the baskets look like. Only 10 randomly selected rows are displayed. Items_bought column shows the baskets.

```{r}


dat2[, items_bought := unlist(list_items) ]

head(dat2[sample(1:nrow(dat2), 10),
          .(Transaction, number_items,items_bought)]) %>%
  data_table()

```

A small example which I will work out manually to see what is the support for ({coffee, bread} --\> {jam}). Generally I want to see how many transactions contained these 3 items. 0.12% of the transactions contained {bread, coffee, jam}

```{r}


my_item_set <- c("Coffee", "Jam", "Bread")

idx_sample <- grep( "Transactio|^Coffee$|^Jam$|^Bread$", names(dat2))



item_set_dat <- dat2[, idx_sample, with = F] 

#some transaction bought more thanone of 
item_set_dat[, (my_item_set) := lapply(.SD, function(x) ifelse(x==0, 0, 1)), .SDcols = my_item_set]

item_set_dat[, total_items := rowSums(item_set_dat[, 2:4, with = F]) ]

support_coffee_bread_jam <- table(item_set_dat$total_items)["3"]/9531 

support_coffee_bread_jam
```

To calculate confidence({bread, coffee} --\> {jam}) we should also calculate the support of ({bread, coffee}) which is the prorpotion of bread and coffee appearing together in the transactions which is about 8.9% of the transactions.

```{r}
item_set_dat[, coffee_bread := rowSums(item_set_dat[, c("Bread", "Coffee"), with = F]) ]

head(item_set_dat, 2)
support_coffee_bread <- table(item_set_dat$coffee_bread)["2"]/9531 

support_coffee_bread
```

1.3% of the people who bought bread and coffee also bought jam. This is the confidence of({bread, coffee} --\> {jam}) For statisticians this can be translated as conditional probability. In conditional probability notations P(Jam/bread, coffee) which is probability you will buy jam given that you have already bought bread and coffee. In association analysis we have {bread, coffee} \>\>{jam} bread and coffee implies jam. So the confidence measures the strength/probability of this implication.

```{r}
confidence <- support_coffee_bread_jam/support_coffee_bread 

confidence * 100
```

Using package arules we find the 10 rules with the highest confidence in descending order. Confidence({Toast} --\>{Coffee}) had the highest confidence of 0.70440252. About 70.44% of the transactions that contained toast also contained coffee.

```{r}
library(arules)

transactions <- as(split(dat$Item, dat$Transaction), "transactions")

assoc_rules <- apriori( transactions, parameter = list(supp = 0.02, conf = 0.04, target = "rules") )

assoc_rules <- sort(assoc_rules, by = "confidence", decreasing = TRUE) 
inspect(assoc_rules[1:30])

```

I hope with this small example you can now understand how association analysis works.