---
title: "Bootstrap in Practice: CIs, Hypothesis Tests, and a Crab-Data Case Study"
format: html
date: "2026-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(broom)
library(kableExtra)
library(glm2)
library(data.table)
data(crabs)
crabs <- na.omit(crabs)

set.seed(12345)
B <- 2000

percentile_ci <- function(x, level = 0.95) {
  alpha <- (1 - level) / 2
  qs <- unname(quantile(x, probs = c(alpha, 1 - alpha), na.rm = TRUE))
  paste0("[", paste(round(qs, 3), collapse = ", "), "]")
}

stat_with_ci <- function(x, stat = mean, level = 0.95) {
  stat <- match.fun(stat)
  est <- stat(x, na.rm = TRUE)
  paste0(round(est, 3), " ", percentile_ci(x, level = level))
}

bootstrap_p_value <- function(t_boot, t_obs, two_sided = TRUE) {
  if (two_sided) {
    (1 + sum(abs(t_boot) >= abs(t_obs))) / (length(t_boot) + 1)
  } else {
    (1 + sum(t_boot >= t_obs)) / (length(t_boot) + 1)
  }
}

options(scipen = 999)

library(DT)
data_table <- function(df, pageLength = 50){
  n_col = ncol(df)-1
  datatable(df,
            rownames = FALSE,
            style = "bootstrap4", class = 'cell-border stripe',
            options = list(scrollX = TRUE,
                           pageLength = pageLength,
                           columnDefs = list(list(className = 'dt-center', targets = 0:n_col))
            )
  )
}
```

## 1. Introduction: what bootstrap is

Any statistic we compute from a sample (a mean, a median, a regression coefficient, a predicted value) has sampling uncertainty: if we repeated the data collection, we would get a slightly different estimate.

The bootstrap is a general, simulation-based way to measure that uncertainty.

-   Treat the observed data as our best stand-in for the population.
-   Generate many new datasets from it.
-   Recompute the statistic on each dataset.
-   Use the spread of those bootstrap replicates to approximate uncertainty.


## 2. The two most common bootstrap types

### 2.1 Nonparametric bootstrap (resample the data)

Resample the observed data with replacement. For i.i.d. data, that usually means resampling rows.

-   Makes minimal distributional assumptions.
-   Very broadly applicable.
-   For hypothesis tests, you still need a way to generate samples under the null (often via centering/transformations, or residual-based bootstraps).

### 2.2 Parametric bootstrap (simulate from a fitted model)

Assume a parametric model is a reasonable approximation, fit it, then simulate new datasets from the fitted model.

-   Stronger assumptions.
-   Particularly natural for likelihood-based models and likelihood-ratio tests.
-   Often excellent for generating null distributions during hypothesis testing, because you can simulate directly under $H_0$. A null distribution is the assumption that the model is correct under $H_0$, so simulating from it gives you a direct way to get a null distribution for your test statistic. eg when doing an independent two-sample t-test, you can simulate a new data set using $N(\hat\mu, \hat\sigma^2)$ under $H_0$.

## 3. Constructing bootstrap estimates and confidence intervals

Let $T$ be a statistic (e.g., $T(x) = \bar x$). The bootstrap procedure is:

1.  Compute $T_{obs} = T(x)$.
2.  For $b=1,\dots,B$:
    -   generate a bootstrap dataset $x^{*(b)}$,
    -   compute $T^{*(b)} = T(x^{*(b)})$.
3.  Use $\{T^{*(b)}\}$ to estimate SEs and CIs.

### 3.1 Nonparametric bootstrap for a mean (and a median)

```{r}
#| label: tbl-bootstrap-mean-median
#| tbl-cap: "Bootstrap estimates for the mean and median of the crab widths, with nonparametric resampling."
set.seed(1)

x <- c(12.1, 10.4, 11.7, 9.9, 10.8, 13.2, 11.0, 10.6, 12.5, 9.8)
n <- length(x)

t_obs_mean <- mean(x)
t_obs_median <- median(x)

boot_mean_np <- numeric(B)
boot_median_np <- numeric(B)

for (b in seq_len(B)) {
    x_star <- sample(x, size = n, replace = TRUE)
    boot_mean_np[b] <- mean(x_star)
    boot_median_np[b] <- median(x_star)
}

se_mean_np <- sd(boot_mean_np)
bias_mean_np <- mean(boot_mean_np) - t_obs_mean

ci_mean_np <- percentile_ci(boot_mean_np)
ci_median_np <- percentile_ci(boot_median_np)

stats_df <- data.table(
    "Mean" = t_obs_mean,
  "Bootstrap Mean" = stat_with_ci(boot_mean_np, stat = mean),
    "Bootstrap SE (Mean)" = se_mean_np,
    "Bootstrap Bias (Mean)" = bias_mean_np,
    "Median" = t_obs_median,
  "Bootstrap Median" = stat_with_ci(boot_median_np, stat = mean)
)

# round only nums
stats_df <- stats_df %>%
    mutate(across(where(is.numeric), ~ round(., 3)))

# melt
stats_df <- melt(stats_df,
    measure.vars = names(stats_df),
    variable.name = "statistic",
    value.name = "value"
)
## put this in a table

data_table(stats_df, pageLength = 10)
```


- The observed sample mean is $\bar x = $ `r round(t_obs_mean, 2)` and the observed median is `r round(t_obs_median, 2)`.
- The bootstrap SE (`mean_boot_se`) is the typical sampling fluctuation of the mean across repeated samples of size `r n`.
- The bootstrap bias estimate (`mean_boot_bias`) is `r round(bias_mean_np, 3)`. With finite $B$, this is a noisy estimate; if it’s small relative to the SE, you typically wouldn’t worry about bias.
- The 95% percentile bootstrap CI for the mean is `r ci_mean_np`. Interpreted literally: if we repeated the whole sampling process many times, about 95% of such intervals would contain the true population mean.
- The 95% percentile bootstrap CI for the median is `r ci_median_np`. Comparing the widths of the mean vs median CIs gives you a feel for which statistic is more variable for this sample size.

### 3.2 Parametric bootstrap for a mean (Normal model)

If we’re comfortable assuming the data are approximately Normal, we can bootstrap by simulating $x^* \sim \mathcal{N}(\hat\mu, \hat\sigma)$.

```{r}
#| label: tbl-bootstrap-mean-parametric
#| tbl-cap: "Bootstrap estimates for the mean of the crab widths, with parametric resampling under a Normal model."
mu_hat <- mean(x)
sigma_hat <- sd(x)

boot_mean_par <- numeric(B)
for (b in seq_len(B)) {
  x_star <- rnorm(n, mean = mu_hat, sd = sigma_hat)
  boot_mean_par[b] <- mean(x_star)
}

se_mean_par <- sd(boot_mean_par) |> round(3)
ci_mean_par <- percentile_ci(boot_mean_par) 

data.table(
  `Bootstrap SE (mean)` = se_mean_par,
  `Bootstrap 95% CI (mean)` = ci_mean_par
) |>
data_table()
```

How to interpret this output

- This CI answers the same question as the nonparametric CI, but *under the Normal model assumption*.
- The parametric bootstrap 95% CI for the mean is `r ci_mean_par`
- If the parametric and nonparametric CIs are very similar here, it suggests the Normal approximation is not doing something wildly different from what the data-driven resampling is doing (at least for the mean).

## 4. Bootstrap hypothesis testing

Bootstrap testing needs one extra ingredient: bootstrap samples must reflect $H_0$.

General recipe:

1.  Choose a test statistic $T$.
2.  Compute $T_{obs}$.
3.  Generate many bootstrap datasets consistent with $H_0$.
4.  Compute $T^{*(b)}$ each time.
5.  Approximate the p-value by how often $T^{*(b)}$ is as extreme as $T_{obs}$.

In the lecture notes, this is called a Monte Carlo p-value. The standard small-sample correction is to use

$$
\hat p = \frac{1 + \#\{T^* \ge T_{obs}\}}{B+1},
$$

and for a two-sided test replace $T$ by $|T|$.

### 4.1 One-sample mean: bootstrap version of a $t$ test

We test $H_0: \mu = \mu_0$ using the usual t-statistic: $$T = \frac{\bar x - \mu_0}{s/\sqrt{n}}.$$

We’ll compare:

-   classical `t.test()`,
-   a nonparametric bootstrap test under $H_0$ (by *recentering* the data),
-   a parametric bootstrap test under $H_0$ (simulate Normal samples with mean $\mu_0$).

These two bootstraps differ only in how they generate samples under $H_0$:

- Nonparametric under $H_0$: recenter the data so it has mean $\mu_0$, then resample with replacement.
- Parametric under $H_0$: assume $X \sim \mathcal{N}(\mu, \sigma^2)$, estimate $\sigma$ from the data, then simulate samples from $\mathcal{N}(\mu_0, \hat\sigma^2)$.

```{r}
mu0 <- 11

t_classical <- t.test(x, mu = mu0)
p_classical <- unname(t_classical$p.value)

# Observed test statistic
t_obs <- unname(t_classical$statistic)

# Nonparametric bootstrap under H0: recenter then resample
x0 <- x - mean(x) + mu0
t_boot_np <- numeric(B)
for (b in seq_len(B)) {
  xs <- sample(x0, size = n, replace = TRUE)
  t_boot_np[b] <- unname(t.test(xs, mu = mu0)$statistic)
}
p_np <- bootstrap_p_value(t_boot_np, t_obs, two_sided = TRUE)

# Parametric bootstrap under H0: simulate Normal with mean mu0
sigma0_hat <- sqrt(var(x))
t_boot_par <- numeric(B)
for (b in seq_len(B)) {
  xs <- rnorm(n, mean = mu0, sd = sigma0_hat)
  t_boot_par[b] <- unname(t.test(xs, mu = mu0)$statistic)
}
p_par <- bootstrap_p_value(t_boot_par, t_obs, two_sided = TRUE)

c(
  classical_t_test = p_classical,
  bootstrap_nonparametric = p_np,
  bootstrap_parametric = p_par
)
```

How to interpret this output

- All three numbers are p-values for the same null, $H_0: \mu = $ `r mu0`, using the same $t$-statistic.
- The observed test statistic is $t_{obs} = $ `r round(t_obs, 3)`. The bootstrap p-values estimate $\Pr(|T^*| \ge |t_{obs}| \mid H_0)$ by simulation.
- If the p-value is small (commonly < 0.05), the observed statistic would be rare under $H_0$, so you have evidence against $H_0$.
- Don’t expect the three p-values to be identical: they differ in how they generate data under $H_0$ (exact $t$ reference vs recenter-and-resample vs Normal simulation).
- Because the bootstrap p-values are Monte Carlo estimates, they have simulation noise. With `r B` replicates, differences in the 3rd decimal place are usually not meaningful.

## 5. Advanced bootstrap: likelihood-ratio testing on crab data

This section uses the horseshoe crab dataset in `glm2` (a classic example for count regression). Reference: [Horseshoe crab dataset in glm2](https://search.r-project.org/CRAN/refmans/glm2/html/crabs.html)

```{r}
#| label: tbl-crabs-head
#| tbl-cap: "First few rows of the horseshoe crab dataset."
#| 
data(crabs)

head(crabs) |>
  data_table()
```



In this study, each row corresponds to a female horseshoe crab. The main variables we use are:

- `Satellites`: the number of satellite males observed with that female (a count outcome, often with many zeros),
- `Width`: the female’s carapace width (a size measure),
- `Dark`: an indicator of shell colour (dark vs not).

The scientific question is: after accounting for size (`Width`), does shell colour (`Dark`) still help explain how many satellites a female has?

The outcome (`Satellites`) is a count, so we fit a Poisson GLM:

-   Reduced (null) model: `Satellites ~ Width`
-   Full model: `Satellites ~ Width + Dark`

### Hypotheses (in words)

- Null hypothesis ($H_0$): after adjusting for female size (`Width`), shell colour (`Dark`) has no association with the expected number of satellite males. In other words, adding `Dark` does not improve the model.
- Alternative hypothesis ($H_1$): after adjusting for `Width`, shell colour (`Dark`) is associated with the expected number of satellite males (so adding `Dark` improves the model).

In model terms, this is the same as testing whether the `Dark` coefficient is zero (for the appropriate coding of `Dark`), typically written as $H_0: \beta_{Dark} = 0$ versus $H_1: \beta_{Dark} \ne 0$.

### 5.1 Classical LRT (chi-square reference)

```{r}
fit_full <- glm(
  Satellites ~ Width + Dark,
  family = poisson(link = "log"),
  data = crabs
)

fit_red <- glm(
  Satellites ~ Width,
  family = poisson(link = "log"),
  data = crabs
)

lrt_table <- anova(fit_red, fit_full, test = "Chisq") |> tidy()
T_obs <- lrt_table$deviance[2]
p_lrt_classical <- lrt_table$p.value[2]
```

```{r}
#| label: tbl-crabs-lrt
#| tbl-cap: "Classical likelihood ratio test (chi-square reference)."
#| tbl-pos: "H"

lrt_table |>
  kable(
    format = "html",
    booktabs = TRUE,
    align = "lccccc",
    escape = FALSE,
    col.names = c("Model", "Resid. Df", "Resid. Dev.", "Df", "$\\chi^2$", "$p$-value")
  ) |>
  kable_styling(latex_options = c("scale_down", "hold_position"),
                font_size = 9, full_width = TRUE) |>
  column_spec(1, width = "5cm")
```

How to interpret this table

- The LRT statistic is $T = 2(\ell_{full} - \ell_{red})$, reported here as `$\\chi^2$`. For these data, $T_{obs} = $ `r round(T_obs, 3)`.
- The chi-square p-value (`r signif(p_lrt_classical, 3)`) is based on a large-sample reference distribution ($T \overset{approx}{\sim} \chi^2_{df}$).
- If this p-value is small, it suggests the full model fits meaningfully better, i.e., `Dark` adds information beyond `Width`.

### 5.2 Parametric bootstrap LRT p-value (null simulation)

Generate bootstrap datasets under $H_0$ by simulating new counts from the reduced model.

Assumptions under $H_0$ (parametric bootstrap):

- The reduced Poisson GLM is correctly specified under the null: conditional on `Width`, satellite counts follow a Poisson distribution with mean $\mu_i$ and log link.
- Observations are independent (each crab is an independent sampling unit).
- The covariates (`Width`, `Dark`) are treated as fixed (or we condition on them). Under $H_0$ the distribution of $Y$ depends on `Width` but not on `Dark`.

This is not a permutation/exchangeability argument: we are not shuffling `Dark` labels. We are simulating $Y$ from an assumed model $Y \mid X \sim \text{Poisson}(\hat\mu_0(X))$.

```{r}
n_crabs <- nrow(crabs)
mu_hat0 <- fitted(fit_red)

T_boot_parametric <- numeric(B)
for (b in seq_len(B)) {
  y_star <- rpois(n_crabs, lambda = mu_hat0)
  dat_star <- crabs
  dat_star$Satellites <- y_star

  fit_red_b <- glm(Satellites ~ Width, family = poisson(link = "log"), data = dat_star)
  fit_full_b <- glm(Satellites ~ Width + Dark, family = poisson(link = "log"), data = dat_star)

  T_boot_parametric[b] <- as.numeric(2 * (logLik(fit_full_b) - logLik(fit_red_b)))
}

p_lrt_parametric <- bootstrap_p_value(T_boot_parametric, T_obs, two_sided = FALSE)
p_lrt_parametric
```

How to interpret this result

- This p-value estimates $\Pr(T^* \ge T_{obs} \mid H_0)$ where $T^*$ is generated by simulating new outcomes from the reduced (null) Poisson GLM.
- Here, $\hat p \approx$ `r signif(p_lrt_parametric, 3)`. If it’s close to the chi-square p-value, that suggests the asymptotic reference is adequate.
- If it differs noticeably, the bootstrap is warning you that the chi-square approximation may be a bit off for this sample/model.

### 5.3 Permutation test (shuffling `Dark`, keeping `Y` and `Width` fixed)

If you want a fully nonparametric way to get a null distribution for the `Dark` effect, a natural option is a permutation test.

Idea: under $H_0$ (`Dark` has no association with satellites after adjusting for width), the `Dark` labels should be exchangeable *given `Width`*.
To make given `Width` more concrete, we do a stratified permutation: bin `Width` and only shuffle `Dark` within bins.

```{r}
set.seed(123)

width_bin <- ntile(crabs$Width, 5)

T_boot_perm <- numeric(B)

for (b in seq_len(B)) {
  dat_perm <- crabs
  dark_perm <- dat_perm$Dark

  for (g in sort(unique(width_bin))) {
    idx_g <- which(width_bin == g)
    dark_perm[idx_g] <- sample(dark_perm[idx_g], size = length(idx_g), replace = FALSE)
  }

  dat_perm$Dark <- dark_perm

  fit_red_b <- glm(Satellites ~ Width, family = poisson(link = "log"), data = dat_perm)
  fit_full_b <- glm(Satellites ~ Width + Dark, family = poisson(link = "log"), data = dat_perm)

  T_boot_perm[b] <- as.numeric(2 * (logLik(fit_full_b) - logLik(fit_red_b)))
}

p_perm <- bootstrap_p_value(T_boot_perm, T_obs, two_sided = FALSE)
p_perm
```

How to interpret this result

- This p-value answers a different “null world” than the parametric bootstrap: it treats the outcome model as unknown and instead tests whether the `Dark` labels are (approximately) exchangeable within `Width` bins under $H_0$.
- Here, $\hat p \approx$ `r signif(p_perm, 3)`. If this is small, then shuffling `Dark` (while preserving the `Width` structure) rarely produces an LRT statistic as large as the observed one.
- If the permutation and parametric-bootstrap p-values disagree, that’s a signal to think carefully about which assumptions are more defensible for this dataset.

```{r}
#| label: fig-lrt-null
#| fig-cap: "Null distributions for the LRT statistic (parametric-bootstrap null and stratified permutation null), with the observed LRT marked."
#| fig-width: 7
#| fig-height: 4

df_plot <- tibble(
  T = c(T_boot_parametric, T_boot_perm),
  method = rep(c("Parametric null", "Permutation null"), each = B)
)

ggplot(df_plot, aes(x = T)) +
  geom_histogram(bins = 40, color = "white") +
  geom_vline(xintercept = T_obs, linewidth = 1) +
  facet_wrap(~ method, ncol = 1, scales = "free_y") +
  labs(x = "LRT statistic", y = "Count") +
  theme_minimal()
```

## 6. Prediction intervals with bootstrap

Confidence intervals quantify uncertainty for an unknown parameter (or mean prediction). Prediction intervals are wider because they must also include the randomness of a *new observation*.

We’ll build a bootstrap prediction interval for the number of satellites for a new crab with a chosen width and shell color.

-   CI for the mean: uncertainty in $\mu(x)$.
-   Prediction interval: uncertainty in a future $Y$ at that $x$ (mean uncertainty + Poisson noise).

```{r}
new_crab <- tibble(Width = 26, Dark = factor("yes", levels = levels(crabs$Dark)))

pred_mean_boot <- numeric(B)
pred_y_boot <- numeric(B)

for (b in seq_len(B)) {
  idx <- sample.int(n_crabs, n_crabs, replace = TRUE)
  dat_b <- crabs[idx, ]

  fit_b <- glm(Satellites ~ Width + Dark, family = poisson(link = "log"), data = dat_b)
  mu_b <- as.numeric(predict(fit_b, newdata = new_crab, type = "response"))

  pred_mean_boot[b] <- mu_b
  pred_y_boot[b] <- rpois(1, lambda = mu_b)
}

ci_mean <- percentile_ci(pred_mean_boot)
pi_y <- percentile_ci(pred_y_boot)

list(
  mean_CI_95 = ci_mean,
  prediction_interval_95 = pi_y
)
```

How to interpret this output

- The 95% CI for the mean (`mean_CI_95`) is for the *expected* number of satellites for a crab with `Width = 26` and `Dark = yes`.
- The 95% prediction interval (`prediction_interval_95`) is for a *single future crab’s observed count* at that same $(Width, Dark)$. It is wider because it includes both parameter uncertainty and Poisson outcome noise.
- If you care about the average response, use the CI; if you care about what an individual new observation might look like, use the prediction interval.

## 7. Conclusion

-   Bootstrap is a general-purpose tool for uncertainty when analytic formulas are fragile or inconvenient.
-   Nonparametric bootstrap resamples the observed data; parametric bootstrap simulates from a fitted model.
-   For confidence intervals, percentile bootstrap is a simple baseline; for hypothesis tests, bootstrap samples must reflect $H_0$.
-   For GLMs and likelihood ratios, parametric (null) bootstrap is a natural way to avoid relying only on asymptotic chi-square approximations.
-   Prediction intervals via bootstrap are often the most directly useful output because they live on the outcome scale.