---
title: "Writing Efficient R Code"
author: "Mburu"
date: "8/13/2021"
output: 
    github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## R version

One of the relatively easy optimizations available is to use an up-to-date version of R. In general, R is very conservative, so upgrading doesn't break existing code. However, a new version will often provide free speed boosts for key functions.

The version command returns a list that contains (among other things) the major and minor version of R currently being used.

```{r}
# Print the R version details using version
version

# Assign the variable major to the major component
major <- version$major

# Assign the variable minor to the minor component
minor <- version$minor
```

## Comparing read times of CSV and RDS files

One of the most common tasks we perform is reading in data from CSV files. However, for large CSV files this can be slow. One neat trick is to read in the data and save as an R binary file (rds) using saveRDS(). To read in the rds file, we use readRDS().

Note: Since rds is R's native format for storing single objects, you have not introduced any third-party dependencies that may change in the future.

To benchmark the two approaches, you can use system.time(). This function returns the time taken to evaluate any R expression. For example, to time how long it takes to calculate the square root of the numbers from one to ten million, you would write the following:

```{r}

# How long does it take to read movies from CSV?
system.time(read.csv("movies.csv"))

# How long does it take to read movies from RDS?
system.time(readRDS("movies.rds"))

```

## Elapsed time

Using system.time() is convenient, but it does have its drawbacks when comparing multiple function calls. The microbenchmark package solves this problem with the microbenchmark() function.

```{r}
# Load the microbenchmark package
library(microbenchmark)

# Compare the two functions
compare <- microbenchmark(read.csv("movies.csv"), 
                          readRDS("movies.rds"), 
                          times = 100)

# Print compare
compare
```

My hardware For many problems your time is the expensive part. If having a faster computer makes you more productive, it can be cost effective to buy one. However, before you splash out on new toys for yourself, your boss/partner may want to see some numbers to justify the expense. Measuring the performance of your computer is called benchmarking, and you can do that with the benchmarkme package.

```{r}
# Load the benchmarkme package
library(benchmarkme)

# Assign the variable ram to the amount of RAM on this machine
ram <- get_ram()
ram

# Assign the variable cpu to the cpu specs
cpu <- get_cpu()
cpu
```

## Benchmark DataCamp's machine

The benchmarkme package allows you to run a set of standardized benchmarks and compare your results to other users. One set of benchmarks tests is reading and writing speeds.

The function call

res = benchmark_io(runs = 1, size = 5) records the length of time it takes to read and write a 5MB file.

```{r, message = TRUE}

# Run the io benchmark
res <- benchmark_io(runs = 1, size = 50)

# Plot the results
plot(res)
```

## Benchmark r operations

```{r, message = TRUE}
# Run each benchmark 3 times
res <- benchmark_std(runs = 10)
plot(res)
```

## Timings - growing a vector

Growing a vector is one of the deadly sins in R; you should always avoid it.

The growing() function defined below generates n random standard normal numbers, but grows the size of the vector each time an element is added!

Note: Standard normal numbers are numbers drawn from a normal distribution with mean 0 and standard deviation 1.

n \<- 30000 \# Slow code growing \<- function(n) { x \<- NULL for(i in 1:n) x \<- c(x, rnorm(1)) x }

```{r}
growing <- function(n) {
    x = NULL
    for(i in 1:n) 
        x = c(x, rnorm(1))
    x
}

# Use <- with system.time() to store the result as res_grow
system.time(res_grow <- growing(30000))
```

Timings - pre-allocation In the previous exercise, growing the vector took around 2 seconds. How long does it take when we pre-allocate the vector? The pre_allocate() function is defined below.

```{r}
n <- 30000
# Fast code
pre_allocate <- function(n) {
    x <- numeric(n) # Pre-allocate
    for(i in 1:n) 
        x[i] <- rnorm(1)
    x
}


# Use <- with system.time() to store the result as res_allocate
n <- 30000
system.time(res_allocate <- pre_allocate(n))
```

## Vectorized code: multiplication

The following piece of code is written like traditional C or Fortran code. Instead of using the vectorized version of multiplication, it uses a for loop.

Your job is to make this code more "R-like" by vectorizing it.

```{r}
x <- rnorm(10)
x2 <- numeric(length(x))
for(i in 1:10)
    x2[i] <- x[i] * x[i]
# Store your answer as x2_imp
x2_imp <- x*x
```

## Vectorized code: calculating a log-sum

A common operation in statistics is to calculate the sum of log probabilities. The following code calculates the log-sum (the sum of the logs).

# x is a vector of probabilities

total \<- 0 for(i in seq_along(x)) total \<- total + log(x\[i\]) However this piece of code could be significantly improved using vectorized code.

```{r}
# Initial code
n <- 100
total <- 0
x <- runif(n)
for(i in 1:n) 
    total <- total + log(x[i])

# Rewrite in a single line. Store the result in log_sum
log_sum <- sum(log(x))
```

## Data frames and matrices - column selection

All values in a matrix must have the same data type, which has efficiency implications when selecting rows and columns.

Suppose we have two objects, mat (a matrix) and df (a data frame).

```{r, eval=FALSE}
# Which is faster, mat[, 1] or df[, 1]? 
microbenchmark(mat[, 1], df[, 1])
```

## Row timings

Similar to the previous example, the objects mat and df are a matrix and a data frame, respectively. Using microbenchmark(), how long does it take to select the first row from each of these objects?

To make the comparison fair, just use mat\[1, \] and df\[1, \].

-   Interesting! Accessing a row of a data frame is much slower than accessing that of a matrix, more so than when accessing a column from each data type. This is because the values of a column of a data frame must be the same data type, whereas that of a row doesn't have to be. Do you see the pattern here?

```{r, eval=FALSE}
# Which is faster, mat[, 1] or df[, 1]? 
microbenchmark(mat[1, ], df[1, ])
```

## Profvis in action

Examine the code on the right that performs a standard data analysis. It loads and selects data, plots the data of interest, and adds in a regression line.

```{r}
# Load the data set
data(movies, package = "ggplot2") 

# Load the profvis package
library(profvis)

# Profile the following code with the profvis function
profvis({
  # Load and select data
  comedies <- movies[movies$Comedy == 1, ]

  # Plot data of interest
  plot(comedies$year, comedies$rating)

  # Loess regression line
  model <- loess(rating ~ year, data = comedies)
  j <- order(comedies$year)
  
  # Add fitted line to the plot
  lines(comedies$year[j], model$fitted[j], col = "red")
})    ## Remember the closing brackets!
```
