---
title: "Heart Disease Explainable ML"
author: "Mburu"
date: "7/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
```

## Heart Disease data set
correlation network plot ml
In health research we are not only interested with prediction level but explanability of the fitted algorithm what people in health research call risk factors. We can use  [heart disease data set](https://www.kaggle.com/ronitf/heart-disease-uci) to figure out how we can utilize some of the mc.

I have stolen your idea. For someone working in health research your notebook is a gold mine. In health especially cohort studies, clinical trials people are largely still using generalized linear models because at least from them you can get odds/risk/rate ratios, p values, AIC and so on. I have implemented the same with R(see here). I have also searched for other materials online see here. Thank you.

```{r cars}
library(tidyverse)
library(data.table)
library(caret)
library(DT)
library(iml)
library(gridExtra)
heart <- fread("heart.csv")

heart %>% head %>%
    datatable(options = list(scrollX = TRUE))


```


## Some Cleaning

You can can check the variable levels from [here](https://archive.ics.uci.edu/ml/datasets/Heart+Disease)

```{r}
heart[, sex := factor(sex, 
                      levels = 0:1, 
                      labels = c("female", "male"))]


heart[, cp := factor(cp,
                     levels = 0:3,
                     labels = c("angina", "atypical angina",
                                "non-anginal pain", "asymptomatic") )]

heart[, fbs := factor(fbs, levels = 0:1,
                      labels = c("fasting blood sugar <= 120 mg/dl)",
                                 "fasting blood sugar > 120 mg/dl)"))]


lnls_restg <- c("normal", 
                 "having ST-T wave abnormality \n (T wave inversions and/or \n ST elevation or depression
                of >0.05 mV)", 
                 "showing probable or definite \n left ventricular hypertrophy \n by Estes' criteria")
 
heart[, restecg := factor(restecg,
                          levels = 0:2,
                          labels = lnls_restg)]



heart[, exang := factor(exang, 
                        levels = 0:1,
                        labels = c("yes", "no"))]


 
heart[, slope := factor(slope,
                         levels = 0:2,
                         labels = c("upsloping", "flat", "downsloping"))]
 
heart[thal == 0, thal := 1 ]
heart[, thal := factor(thal,
                        levels =  1:3,
                        labels = c("normal", "fixed defect", "reversable defect"))]

heart[, ca := factor(ca)]

heart[, target := factor(target,
                         levels = 0:1,
                         labels = c("No_heart_disease", "Heart_disease"))]

```


## Visualize categorical variables

```{r, fig.width =15, fig.height = 20}
nms <- names(heart)
categ_nms <- nms[sapply(heart, is.factor)]


categ_nms <- categ_nms[categ_nms != "target"]
plots_categ <- list()
for (i in categ_nms) {
    
   plots_categ[[i]] = heart[, .(freq = .N), by = c(i,"target")] %>%
        .[, perc := round(freq/sum(freq) * 100, 2), by = i] %>%
        ggplot(aes_string(i, "perc", fill = "target")) +
        geom_bar(stat = "identity", width = 0.5)+
       geom_text(aes_string(i, "perc", label = "perc"),
                 size = 3 , position =  position_stack(vjust = 0.5))+
       theme(legend.position = "bottom")+
       scale_fill_brewer(palette  = "Dark2")

}

grid.arrange(grobs = plots_categ, ncol = 2)
```


## Visualize numeric variable

```{r, fig.width = 7, fig.height = 4}
num_nms <- nms[sapply(heart, is.numeric)]

# zero_one <- function(x){
#   minx = min(x, na.rm = T)
#   maxx = max(x, na.rm = T)
#   
#   z = (x - minx)/(maxx - minx)
# }
# 
# heart[, (num_nms) := lapply(.SD, zero_one), .SDcols = num_nms]

num_nms <- num_nms[num_nms != "target"]

for (i in num_nms) {
    
        p = ggplot(heart, aes_string("target", i)) +
            geom_boxplot()
   print(p)
}
```


## Fit logistic regression

```{r}
glm_heart <- glm(target ~., data = heart,
                 family = binomial())

library(broom)

tidy(glm_heart) %>% datatable() %>%
  formatRound(columns = c("estimate", "std.error", "statistic", "p.value"), digits = 4)
```


## Chisq test


```{r}
drop1(glm_heart, test = "Chisq") %>% tidy %>%
    datatable()  %>%
  formatRound(columns = c("df", "Deviance", "AIC", "LRT", "p.value"), digits = 4)
```


## ML model

```{r}

train_ctrl <- trainControl(method = "cv",
                        number = 5,
                        summaryFunction = twoClassSummary,
                        classProbs = TRUE,
                        allowParallel=T,
                        #index = cv_fold,
                        verboseIter = FALSE,
                        returnResamp = "all", 
                        savePredictions = "final", 
                        search = "grid")





ranger_grid <- expand.grid(splitrule = "extratrees",
                        mtry = c(2, 5, 10),
                        min.node.size = c(2, 5, 7))


heart_randomforest <- train(target~ .,
      data = heart,
      trControl = train_ctrl,
      tuneGrid = ranger_grid,
      method = "ranger")
heart_randomforest 
```
## Var Importance
```{r}
pred <- function(heart_randomforest, heart)  {
  results <- predict(heart_randomforest, newdata = heart, type = "prob")
  return(results[[2L]])
}
```

## Partial Dependence

```{r}
X_pred <- heart[, .SD, .SDcols = !"target"] %>%
  as.data.frame()
model <- Predictor$new(model = heart_randomforest, 
                      data =X_pred,
                      predict.function = pred,
                      y = heart$target)
effect <- FeatureEffects$new(model)

```

##

```{r}
effect$plot(features = c( "trestbps"))

#effect
```



##

```{r}
imp <- FeatureImp$new(model, loss ="ce" )
#imp

var_important <-imp$results %>% data.table()

setorder(var_important, -importance)
plot(imp)
```

```{r}
interact <- Interaction$new(model, feature = "thal")
plot(interact)
```


```{r}
library(tictoc)
tic()

shap_list <- vector("list", nrow(X_pred)) 

for (i in 1:nrow(X_pred)) {
  shap <- Shapley$new(model,  x.interest = X_pred[i, ], sample.size = 30)
  shap_import <-shap$results %>% data.table()
  shap_import <- shap_import[class == "Heart_disease"]
  shap_list[[i]] <- shap_import[, record_id := i]

  }
toc()

shap_values <- rbindlist(shap_list, fill = T)

```



```{r}

library(ggforce)

shap_values[, feature := factor(feature, levels = rev(var_important$feature) )]
ggplot(shap_values, aes(feature, phi,  color = phi.var))+
  geom_sina()+
  geom_hline(yintercept = 0) +
  scale_color_gradient(low="#2187E3", high="#F32858", 
                       breaks=c(0,1), labels=c("Low","High"))+ 
  theme_bw() + 
    theme(axis.line.y = element_blank(), 
          axis.ticks.y = element_blank(), # remove axis line
          legend.position="bottom") +
  coord_flip()

```

