---
title: "Kenya Mobile Phones Usage and Financial Inclusion Insights from FinAccess 2021"
author: "Mburu"
date:  "2024-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 500)

```

```{r, include=FALSE}
source("R/data_check.R")
```


The FinAccess survey's graph presents a vivid portrayal of monthly income distribution across different brackets in Kenya for the year 2021. It's a stark reminder that a substantial portion of the Kenyan workforce operates within the informal sector, often earning a modest income that may not be consistently captured in formal tax structures.

In contemplating tax increases to address fiscal deficits and debt service costs, policymakers must consider the reality depicted by the graph: the majority of Kenyans fall within lower income groups, with a considerable number earning less than KSh 5,000 a month. Imposing higher taxes could disproportionately affect these individuals, diminishing their already limited purchasing power and potentially pushing them into poverty.

Moreover, since the informal sector is not thoroughly integrated into the formal tax system, tax hikes may not significantly increase revenue. Instead, they could drive more economic activity underground, away from the purview of tax authorities, and potentially erode the tax base.


```{r, fig.width=9.5, fig.height=9.5}
# save map 5 as .RDA
fina_inco_map <- map5
save(fina_inco_map, file = "R/fina_inco_map.RDA")
girafe(ggobj = map5)

```

```{r, echo=FALSE,result="hide"}
gc()
```


```{r}
#girafe(ggobj = map3)
```

```{r, fig.width=9.5, fig.height=9.5}
#map1
#girafe(ggobj = map1)
```

```{r, fig.width=9.5, fig.height=9.5}
#girafe(ggobj = map3)
```

```{r, fig.width=9.5, fig.height=9.5}
#girafe(ggobj = map4)

```

