---
title: "GitHub Innovation Graph data"
author: "Mburu"
date: "2024-02-29"
output:
    html_document:
        include:
            in_header: world.html

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)



```

## Introduction

This is a similar analysis to my previous which looked into GitHub use in Africa. I used  [GitHub Innovation Graph data](https://github.com/github/innovationgraph/tree/main) to see how GitHub public repositories activity is distributed around the world.

```{r}
#| code-fold: true
library(tidyverse)
library(data.table)
library(janitor)
library(tmap)
library(patchwork)
library(sf)
library(ggiraph)
library(viridis) # For scale_fill_viridis_c

developers <- fread("https://raw.githubusercontent.com/github/innovationgraph/main/data/developers.csv")
git_pushes <- fread("https://raw.githubusercontent.com/github/innovationgraph/main/data/git_pushes.csv")
git_repos <- fread("https://raw.githubusercontent.com/github/innovationgraph/main/data/repositories.csv")
programming_languages <- fread("https://raw.githubusercontent.com/github/innovationgraph/main/data/languages.csv")

merge_by_devs <- c("iso2_code", "year", "quarter")
merge_fun_devs <- function(devs, df, by.vars = merge_by_devs)  {
    
    merge(devs, df, by = by.vars)
    
}

git_pushes <- merge_fun_devs(developers, git_pushes)
git_repos <- merge_fun_devs(developers, git_repos)
programming_languages <- merge_fun_devs(developers, programming_languages)

github_data_name <- c("developers", "git_pushes", "git_repos", "programming_languages")

iso3 <- fread("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/all/all.csv") 
iso3 <- iso3[, .(iso2_code = `alpha-2`, iso_a3 = `alpha-3`)]

data("World") # from tmap


africa <- World %>% 
    left_join(iso3,by = "iso_a3")# %>%
    #filter(continent == "Africa")


setDT(africa)


get_latest_year_and_quarter <- function(df, year_column, quarter_column, concactenate = TRUE) {
    # Calculate the latest year with available data
    latest_year <- max(df[[year_column]], na.rm = TRUE)
    
    # Calculate the latest quarter for the latest year
    latest_quarter <- max(df[df[[year_column]] == latest_year][[quarter_column]], na.rm = TRUE)
    
    # Check if the user wants to return a concatenated contact string
    if (concactenate) {
        # Return the concatenated string of latest year and quarter
        return(paste0(latest_year, "-", latest_quarter))
    } else {
        # Return a named list containing the latest year and quarter
        return(list(latest_year = latest_year, latest_quarter = latest_quarter))
    }
}


# Define the function
process_df_to_sf <- function(df, spatial_df,
                             iso2_column_name = "iso2_code", 
                             year_column = "year",
                             quarter_column = "quarter",
                             pop_est_column = "pop_est", 
                             freq_column = "developers",
                             geometry_column = "geometry",
                             group_col = NULL,
                             per = 100000,
                             round_digits = 0) {
    # Convert df to a data.table
    setDT(df)
    
    # Filter for the latest year and quarter
    latest_year_quarter<- get_latest_year_and_quarter(df, year_column, quarter_column,  concactenate = FALSE)
    latest_year <-latest_year_quarter$latest_year #  max(df[[year_column]], na.rm = TRUE)
    latest_quarter <- latest_year_quarter$latest_quarter # max(df[get(year_column) == latest_year][[quarter_column]], na.rm = TRUE) get
    #print(latest_year)
    #print(latest_quarter)
    latest_df <- df[get(year_column) == latest_year & get(quarter_column) == latest_quarter]
    # print(head(latest_df))
    # Join with the spatial data frame
    
    if(!is.null(group_col)) {
        latest_df <- latest_df[!is.na(get(group_col)),]
        dfs = split.data.frame(latest_df, f = latest_df[[group_col]])
        #cat("length dfs:", length(dfs))
        
        ## right join by spacial df and rbindlist
        latest_df_joined = rbindlist(lapply(dfs, function(x) {
            uniqu_grp_col = unique(x[[group_col]])
            x = x %>%
                merge(spatial_df, by = iso2_column_name, all.y = TRUE)
            x[, (group_col) :=uniqu_grp_col]
            x
        })) 
        
        #print(head(latest_df_joined))
        
    }else{
        
        latest_df_joined <- latest_df %>%
            merge(spatial_df, by = iso2_column_name, all.y = TRUE)
        
    }
    
    latest_df_joined[is.na(get(freq_column)), (freq_column) := 0]
    #latest_df_joined[is.na(get(pop_est_column)), (pop_est_column) := 0]
    latest_df_joined[, devs_per_100k := round(.SD[[freq_column]] / .SD[[pop_est_column]] * per, digits = round_digits), .SDcols = c(freq_column, pop_est_column)]
    latest_df_joined[is.na(devs_per_100k), devs_per_100k := 0]
    # Ensure spatial_df is a data.frame or data.table with a geometry column
    if(!"geometry" %in% colnames(spatial_df)) {
        stop("The spatial data frame must have a geometry column.")
    }
    
    # Convert to sf object, ensuring the geometry column is specified correctly
    latest_df_sf <- st_set_geometry(latest_df_joined, value = geometry_column)
    
    # Return the sf object
    return(latest_df_sf)
}




# Define the function with tidy evaluation for aes()
create_interactive_map <- function(data_sf, 
                                   fill_var, 
                                   tooltip_var,
                                   facet_var = NULL,
                                   plot_title = "Developers per 100,000 people",
                                   return_girafe = TRUE,
                                   aspect_ratio = 1,
                                   point_size = 9)  {
    # Create the ggplot
    p <- ggplot(data_sf, aes(fill = {{fill_var}}, tooltip = paste0({{tooltip_var}}, ":", {{fill_var}}) )) +
        geom_sf_interactive() +
        scale_fill_viridis_c() +
        theme_void() +
        theme(legend.position = "left",
              legend.key.width = unit(0.1, 'cm'),
              plot.title = element_text(size = 11),
              aspect.ratio = aspect_ratio,
              plot.margin = unit(c(0,0,0,0), "cm"))+
        labs(title = plot_title, fill = "")
    
    # Dynamically add facets if a facet_var is provided
    if (!is.null(facet_var)) {
        # Use the .data pronoun to refer to variables for faceting
        p <- p + facet_wrap(vars(.data[[facet_var]]))
    }
    
    # Generate the interactive map
    if (return_girafe){
        girafe_code <- girafe(ggobj = p, pointsize = point_size)
        return(girafe_code)
    } else {
        return(p)
    }
    
}

# Example usage:
# Assuming 'latest_developers_sf' is your sf dataframe and has 'devs_per_100k' for fill and 'name' for tooltip
# create_interactive_map(latest_developers_sf, fill_var = devs_per_100k, tooltip_var = paste0(name, ":", devs_per_100k))
# To add facets, simply add the facet_var argument like facet_var = "region"



library(ggplot2)
library(ggiraph)
library(stringr) # For str_wrap


plot_interactive_bar <- function(data, x_var, y_var, 
                                 fill_var, 
                                 title = "Top programming languages by country",
                                 xlab = "Country", 
                                 ylab = "% Number of pushers") {
  
  # Create the plot
  top_countries_plot <- ggplot(data, 
                               aes(x = {{x_var}}, y = {{y_var}}, fill = {{fill_var}},
                                   tooltip = paste0({{x_var}}, ", ", {{fill_var}}, " ", {{y_var}}))) +
      
      
    geom_bar_interactive(stat = "identity", position = "dodge", width = 0.5) +
    theme_minimal() +
    labs(title = title,
         fill = "",
         y = ylab,
         x = xlab) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    theme(legend.position = "bottom")
  
  # Return the interactive plot
  girafe(ggobj = top_countries_plot, pointsize = 9)
}


## donhut chart

donhunt_chart <- function(data, labelColumn, valueColumn, chartTitle) {
    plot_ly(data, labels = ~get(labelColumn), values = ~get(valueColumn)) %>%
        add_pie(hole = 0.6) %>%
        layout(title = chartTitle,
               xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
               yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
}


```

## Developers, Repositories, Gitpushes `r  get_latest_year_and_quarter(developers, "year", "quarter")`

### Developers/GitHub users


```{r, fig.width=4, fig.height=4}
latest_developers_sf <- process_df_to_sf(developers, africa,
                                         iso2_column_name = "iso2_code",
                                         freq_column = "developers", 
                                         pop_est_column = "pop_est")

devs_100k <- create_interactive_map(data_sf = latest_developers_sf, 
                                    fill_var = devs_per_100k,
                                    tooltip_var = name,
                                    return_girafe = TRUE,
                                    aspect_ratio = 1,
                                    point_size = 20)
devs_100k
```

### Git Pushes per Developer `r  get_latest_year_and_quarter(git_repos, "year", "quarter")`


```{r, fig.width=4, fig.height=4}
git_pushes_sf <- process_df_to_sf(git_pushes, africa,
                                  iso2_column_name = "iso2_code",
                                  freq_column = "git_pushes", 
                                  pop_est_column = "developers",
                                  per = 1)

github_pushes_map <- create_interactive_map(data_sf = git_pushes_sf, 
                                            fill_var = devs_per_100k,
                                            tooltip_var = name,
                                            plot_title = "Average GitHub pushes per developer",
                                            return_girafe = TRUE,
                                            aspect_ratio = 1,
                                            point_size = 20)
github_pushes_map
```

### GitHub Repositories per Developer `r  get_latest_year_and_quarter(git_repos, "year", "quarter")`


```{r, fig.width=5, fig.height=5}


git_repos_sf <- process_df_to_sf(git_repos, africa,
                                 iso2_column_name = "iso2_code",
                                 freq_column = "repositories", 
                                 pop_est_column = "developers",
                                 per = 1)

github_repos_map <- create_interactive_map(data_sf = git_repos_sf,
                                           fill_var = devs_per_100k,
                                           tooltip_var = name,
                                           plot_title = "Average GitHub repositories  per developer",
                                           return_girafe = TRUE,
                                           aspect_ratio = 1,
                                           point_size = 20)



github_repos_map
```

## Programming languages `r  get_latest_year_and_quarter(programming_languages, "year", "quarter")`

### Top 10 programming languages


-   Top 10 programming languages by the number pushers

```{r, fig.width=5, fig.height=5}
latest_developers_sf_top <- top_n(latest_developers_sf, 10, devs_per_100k)
programming_languages_sf <- process_df_to_sf(programming_languages,
                                             spatial_df = africa,
                                             iso2_column_name = "iso2_code",
                                             freq_column = "num_pushers", 
                                             pop_est_column = "developers",
                                             group_col = "language",
                                             per = 100,
                                             round_digits = 2)
programming_languages_dt <- as.data.table(programming_languages_sf)
programming_languages_dt[, geomety := NULL]
programming_languages_dt[, name := as.character(name)]

## top 10 programming languages by the number of pushers
top_10 <- programming_languages_dt[,
                                  .(num_pushers = sum(num_pushers)),
                                  by = language][order(-num_pushers), .SD[1:15]]

library(plotly)

#plot pie chart for top_10
top_lan <-donhunt_chart(data = top_10, 
                        labelColumn = "language", 
                        valueColumn = "num_pushers", 
                        chartTitle = "Top 15 Programming Languages by number of pushers in the World")

htmltools ::div(top_lan, align = "center") 

```

### Top 3 programming languages

-   Top 3 programming languages by the number pushers in each country
-   Presents only top ten countries by the number of developers per 100,000 people

```{r, fig.width=9, fig.height=5}

number_pushers_lan <- programming_languages_dt[
    iso2_code %in% latest_developers_sf_top$iso2_code]

number_pushers_lan <- number_pushers_lan[devs_per_100k != 0,]

## get top 3 languages with highest number of developers per 100k per country
top3_lan_per_country <- number_pushers_lan[order(-devs_per_100k), .SD[1:3],  by = name]


plot_interactive_bar(data = top3_lan_per_country,
                     x_var = name,
                     y_var = devs_per_100k,
                     fill_var = language,
                     title = "Top 3 languages with highest % of developers that pushed code to GitHub",)


```

### Top 6 Programming Languages in World: Percentage of developers distribution pushing code to GitHub

```{r, fig.width=7, fig.height=7}
# Merge the programming languages data with the spatial data
setorder(top_10, -num_pushers)
programming_languages_sf_top6 <- programming_languages_sf %>%
    filter(language %in% top_10$language[1:6])


create_interactive_map(data_sf = programming_languages_sf_top6,
                       fill_var = devs_per_100k,
                       tooltip_var = name,
                       facet_var = "language",
                       plot_title = "Distribution of top 6 programming languages in World (% of developers that pushed code to GitHub)")

```

### A look at R: Used for analytics and statistics

-   Analytics/Data Science
-   Python presented on the map before this

```{r}
#"R", "Python", "Java",
filter_sf_obj <- function(sf_obj, languages) {
    sf_obj %>% 
        filter(language %in% languages)
}


r_pyh <- filter_sf_obj(programming_languages_sf, 
                       languages = c("R"))

create_interactive_map(data_sf =r_pyh ,
                       fill_var = num_pushers,
                       tooltip_var = name,
                       #facet_var = "language",
                       plot_title = "Distribution of R users in World")

```

### Number of pushers; A look at Java & Swift : Languages as a proxy for mobile app development

```{r}
#"R", "Python", "Java",
java_swift <- filter_sf_obj(programming_languages_sf,
                            languages = c("Java", "Swift"))


create_interactive_map(data_sf =java_swift ,
                       fill_var = num_pushers,
                       tooltip_var = name,
                       facet_var = "language",
                       plot_title = "Countries leading in Java or Swift")

```
