---
title: "Functions in R"
date: "2022-12-10"
output:
  html_document:
    toc: yes
    toc_depth: 3
    theme: lumen
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(data.table)
library(knitr)
```

## Calling functions

One way to make your code more readable is to be careful about the order you pass arguments when you call functions, and whether you pass the arguments by position or by name.

gold_medals, a numeric vector of the number of gold medals won by each country in the 2016 Summer Olympics, is provided.

For convenience, the arguments of median() and rank() are displayed using args(). Setting rank()'s na.last argument to "keep" means "keep the rank of NA values as NA".

Best practice for calling functions is to include them in the order shown by args(), and to only name rare arguments.

```{r}

gd_nms <-  c("USA", "GBR", "CHN", "RUS", "GER", "JPN", "FRA", "KOR", "ITA", 
            "AUS", "NED", "HUN", "BRA", "ESP", "KEN", "JAM", "CRO", "CUB", 
            "NZL", "CAN", "UZB", "KAZ", "COL", "SUI", "IRI", "GRE", "ARG", 
            "DEN", "SWE", "RSA", "UKR", "SRB", "POL", "PRK", "BEL", "THA", 
            "SVK", "GEO", "AZE", "BLR", "TUR", "ARM", "CZE", "ETH", "SLO", 
            "INA", "ROU", "BRN", "VIE", "TPE", "BAH", "IOA", "CIV", "FIJ", 
            "JOR", "KOS", "PUR", "SIN", "TJK", "MAS", "MEX", "VEN", "ALG", 
            "IRL", "LTU", "BUL", "IND", "MGL", "BDI", "GRN", "NIG", "PHI", 
            "QAT", "NOR", "EGY", "TUN", "ISR", "AUT", "DOM", "EST", "FIN", 
            "MAR", "NGR", "POR", "TTO", "UAE", "IOC")

gold_medals <- c(46L, 27L, 26L, 19L, 17L, 12L, 10L, 9L, 8L, 8L, 8L, 8L, 7L, 
                 7L, 6L, 6L, 5L, 5L, 4L, 4L,4L, 3L, 3L, 3L, 3L, 3L, 3L, 2L, 2L, 
                 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 1L,
                 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L,
                 1L, 1L, 1L, 1L, 0L,0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
                 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
                 0L, 0L, 0L, 0L, 0L, 0L, NA)

names(gold_medals) <- gd_nms

gold_medals


# Note the arguments to median()
args(median)

# Rewrite this function call, following best practices
median(gold_medals, na.rm = TRUE)


# Note the arguments to rank()
args(rank)

# Rewrite this function call, following best practices

rank(-gold_medals, na.last = "keep", ties.method=  "min")
```

## Your first function: tossing a coin

Time to write your first function! It's a really good idea when writing functions to start simple. You can always make a function more complicated later if it's really necessary, so let's not worry about arguments for now.

```{r}
coin_sides <- c("head", "tail")

# Sample from coin_sides once
sample(coin_sides, 1)

# Your functions, from previous steps
toss_coin <- function() {
  coin_sides <- c("head", "tail")
  sample(coin_sides, 1)
}

# Call your function
toss_coin()
```

## Inputs to functions

Most functions require some sort of input to determine what to compute. The inputs to functions are called arguments. You specify them inside the parentheses after the word "function."

As mentioned in the video, the following exercises assume that you are using sample() to do random sampling.

```{r}
# Update the function to return n coin tosses
toss_coin <- function(n_flips) {
  coin_sides <- c("head", "tail")
  sample(coin_sides, n_flips, replace = TRUE)
}

# Generate 10 coin tosses
toss_coin(1000) %>% table() %>% prop.table()
```

## Multiple inputs to functions

If a function should have more than one argument, list them in the function signature, separated by commas. To solve this exercise, you need to know how to specify sampling weights to sample(). Set the prob argument to a numeric vector with the same length as x. Each value of prob is the probability of sampling the corresponding element of x, so their values add up to one. In the following example, each sample has a 20% chance of "bat", a 30% chance of "cat" and a 50% chance of "rat".

```{r}
# Update the function so heads have probability p_head
toss_coin <- function(n_flips, p_head) {
  coin_sides <- c("head", "tail")
  # Define a vector of weights
  weights <- c(p_head,1- p_head)
  # Modify the sampling to be weighted
  sample(coin_sides, n_flips, replace = TRUE, prob = weights)
}

# Generate 10 coin tosses
toss_coin(10,p_head=.8)
```

## Data or detail?

Recall that data arguments are what a function computes on, and detail arguments advise on how the computation should be performed. Each of the arguments to t.test() is shown, along with a brief description of it.

## Renaming GLM

R's generalized linear regression function, glm(), suffers the same usability problems as lm(): its name is an acronym, and its formula and data arguments are in the wrong order. To solve this exercise, you need to know two things about generalized linear regression: glm() formulas are specified like lm() formulas: response is on the left, and explanatory variables are added on the right. To model count data, set glm()'s family argument to poisson, making it a Poisson regression. Here you'll use data on the number of yearly visits to Snake River at Jackson Hole, Wyoming, snake_river_visits.

```{r}
snake_river_visits <-read_rds("snake_river_visits.rds")
# From previous step
run_poisson_regression <- function(data, formula) {
  glm(formula, data, family = poisson)
}

# Re-run the Poisson regression, using your function
model <- snake_river_visits %>%
  run_poisson_regression(n_visits ~ gender + income + travel)

# Run this to see the predictions
snake_river_visits %>%
  mutate(predicted_n_visits = predict(model, ., type = "response"))%>%
  arrange(desc(predicted_n_visits)) %>%
    head() %>% kable()
```

## Numeric defaults

cut_by_quantile() converts a numeric vector into a categorical variable where quantiles define the cut points. This is a useful function, but at the moment you have to specify five arguments to make it work. This is too much thinking and typing. By specifying default arguments, you can make it easier to use. Let's start with n, which specifies how many categories to cut x into.A numeric vector of the number of visits to Snake River is provided as n_visits.

```{r}
# Set the default for n to 5
n_visits = snake_river_visits$n_visits
cut_by_quantile <- function(x, n=5, na.rm, labels, interval_type) {
  probs <- seq(0, 1, length.out = n + 1)
  qtiles <- quantile(x, probs, na.rm = na.rm, names = FALSE)
  right <- switch(interval_type, "(lo, hi]" = TRUE, "[lo, hi)" = FALSE)
  cut(x, qtiles, labels = labels, right = right, include.lowest = TRUE)
}

# Remove the n argument from the call
cut_by_quantile(
  n_visits, 
  na.rm = FALSE, 
  labels = c("very low", "low", "medium", "high", "very high"),
  interval_type = "(lo, hi]"
) %>% head()
```

## Logical defaults

cut_by_quantile() is now slightly easier to use, but you still always have to specify the na.rm argument. This removes missing values -- it behaves the same as the na.rm argument to mean() or sd().

Where functions have an argument for removing missing values, the best practice is to not remove them by default (in case you hadn't spotted that you had missing values). That means that the default for na.rm should be FALSE.

```{r}
# Set the default for na.rm to FALSE
cut_by_quantile <- function(x, n = 5, na.rm = FALSE, labels, interval_type) {
  probs <- seq(0, 1, length.out = n + 1)
  qtiles <- quantile(x, probs, na.rm = na.rm, names = FALSE)
  right <- switch(interval_type, "(lo, hi]" = TRUE, "[lo, hi)" = FALSE)
  cut(x, qtiles, labels = labels, right = right, include.lowest = TRUE)
}

# Remove the na.rm argument from the call
cut_by_quantile(
  n_visits, 
  labels = c("very low", "low", "medium", "high", "very high"),
  interval_type = "(lo, hi]"
) %>% head()
```

## NULL defaults

The cut() function used by cut_by_quantile() can automatically provide sensible labels for each category. The code to generate these labels is pretty complicated, so rather than appearing in the function signature directly, its labels argument defaults to NULL, and the calculation details are shown on the ?cut help page.

```{r}
# Set the default for labels to NULL
cut_by_quantile <- function(x, n = 5, na.rm = FALSE, labels=NULL, interval_type) {
  probs <- seq(0, 1, length.out = n + 1)
  qtiles <- quantile(x, probs, na.rm = na.rm, names = FALSE)
  right <- switch(interval_type, "(lo, hi]" = TRUE, "[lo, hi)" = FALSE)
  cut(x, qtiles, labels = labels, right = right, include.lowest = TRUE)
}

# Remove the labels argument from the call
cut_by_quantile(
  n_visits,
  #labels = c("very low", "low", "medium", "high", "very high"),
  interval_type = "(lo, hi]"
) %>% head()
```

## Categorical defaults

When cutting up a numeric vector, you need to worry about what happens if a value lands exactly on a boundary. You can either put this value into a category of the lower interval or the higher interval. That is, you can choose your intervals to include values at the top boundary but not the bottom (in mathematical terminology, "open on the left, closed on the right", or (lo, hi]). Or you can choose the opposite ("closed on the left, open on the right", or [lo, hi)). cut_by_quantile() should allow these two choices.

The pattern for categorical defaults is:

function(cat_arg = c("choice1", "choice2")) { cat_arg \<- match.arg(cat_arg) }

Free hint: In the console, type head(rank) to see the start of rank()'s definition, and look at the ties.method argument.

```{r}
# Set the categories for interval_type to "(lo, hi]" and "[lo, hi)"
cut_by_quantile <- function(x, n = 5, na.rm = FALSE, labels = NULL, 
                            interval_type= c("(lo, hi]", "[lo, hi)")) {
  # Match the interval_type argument
  interval_type <- match.arg(interval_type)
  probs <- seq(0, 1, length.out = n + 1)
  qtiles <- quantile(x, probs, na.rm = na.rm, names = FALSE)
  right <- switch(interval_type, "(lo, hi]" = TRUE, "[lo, hi)" = FALSE)
  cut(x, qtiles, labels = labels, right = right, include.lowest = TRUE)
}

# Remove the interval_type argument from the call
cut_by_quantile(n_visits) %>% head()
```

## Harmonic mean

The harmonic mean is the reciprocal of the arithmetic mean of the reciprocal of the data. That is

The harmonic mean is often used to average ratio data. You'll be using it on the price/earnings ratio of stocks in the Standard and Poor's 500 index, provided as std_and_poor500. Price/earnings ratio is a measure of how expensive a stock is.

The dplyr package is loaded.

```{r}
std_and_poor500 <- read_rds("std_and_poor500_with_pe_2019-06-21.rds")
# Look at the Standard and Poor 500 data
glimpse(std_and_poor500)

# Write a function to calculate the reciprocal
# From previous steps
get_reciprocal <- function(x) {
  1 / x
}
calc_harmonic_mean <- function(x) {
  x %>%
    get_reciprocal() %>%
    mean() %>%
    get_reciprocal()
}

std_and_poor500 %>% 
  # Group by sector
  group_by(sector) %>% 
  # Summarize, calculating harmonic mean of P/E ratio
  summarise(hmean_pe_ratio = calc_harmonic_mean(pe_ratio))
```

## Dealing with missing values

In the last exercise, many sectors had an NA value for the harmonic mean. It would be useful for your function to be able to remove missing values before calculating.

Rather than writing your own code for this, you can outsource this functionality to mean().

The dplyr package is loaded.

```{r}
# From previous step
calc_harmonic_mean <- function(x, na.rm = FALSE) {
  x %>%
    get_reciprocal() %>%
    mean(na.rm = na.rm) %>%
    get_reciprocal()
}

std_and_poor500 %>% 
  # Group by sector
  group_by(sector) %>% 
  # Summarize, calculating harmonic mean of P/E ratio
  summarise(hmean_pe_ratio = calc_harmonic_mean(pe_ratio, na.rm = TRUE))
```

## Passing arguments with ...

Rather than explicitly giving calc_harmonic_mean() and na.rm argument, you can use ... to simply "pass other arguments" to mean().

The dplyr package is loaded.

```{r}
calc_harmonic_mean <- function(x, ...) {
  x %>%
    get_reciprocal() %>%
    mean(...) %>%
    get_reciprocal()
}


std_and_poor500 %>% 
  # Group by sector
  group_by(sector) %>% 
  # Summarize, calculating harmonic mean of P/E ratio
  summarise(hmean_pe_ratio = calc_harmonic_mean(pe_ratio, na.rm = TRUE))
```

## Throwing errors with bad arguments

If a user provides a bad input to a function, the best course of action is to throw an error letting them know. The two rules are

Throw the error message as soon as you realize there is a problem (typically at the start of the function). Make the error message easily understandable. You can use the assert\_\*() functions from assertive to check inputs and throw errors when they fail.

```{r}
library(assertive)
calc_harmonic_mean <- function(x, na.rm = FALSE) {
  # Assert that x is numeric
  assert_is_numeric(x)
  x %>%
    get_reciprocal() %>%
    mean(na.rm = na.rm) %>%
    get_reciprocal()
}

# See what happens when you pass it strings
#calc_harmonic_mean(std_and_poor500$sector)
```

## Custom error logic

Sometimes the assert\_\*() functions in assertive don't give the most informative error message. For example, the assertions that check if a number is in a numeric range will tell the user that a value is out of range, but the won't say why that's a problem. In that case, you can use the is\_\*() functions in conjunction with messages, warnings, or errors to define custom feedback.

The harmonic mean only makes sense when x has all positive values. (Try calculating the harmonic mean of one and minus one to see why.) Make sure your users know this!

```{r}
calc_harmonic_mean <- function(x, na.rm = FALSE) {
  assert_is_numeric(x)
  # Check if any values of x are non-positive
  if(any(is_non_positive(x), na.rm = TRUE)) {
    # Throw an error
    stop("x contains non-positive values, so the harmonic mean makes no sense.")
  }
  x %>%
    get_reciprocal() %>%
    mean(na.rm = na.rm) %>%
    get_reciprocal()
}

# See what happens when you pass it negative numbers
#calc_harmonic_mean(std_and_poor500$pe_ratio - 20)
```

## Fixing function arguments

The harmonic mean function is almost complete. However, you still need to provide some checks on the na.rm argument. This time, rather than throwing errors when the input is in an incorrect form, you are going to try to fix it.

na.rm should be a logical vector with one element (that is, TRUE, or FALSE).

The assertive package is loaded for you.

```{r}
# Update the function definition to fix the na.rm argument
calc_harmonic_mean <- function(x, na.rm = FALSE) {
  assert_is_numeric(x)
  if(any(is_non_positive(x), na.rm = TRUE)) {
    stop("x contains non-positive values, so the harmonic mean makes no sense.")
  }
  # Use the first value of na.rm, and coerce to logical
  na.rm <- coerce_to(use_first(na.rm), target_class = "logical")
  x %>%
    get_reciprocal() %>%
    mean(na.rm = na.rm) %>%
    get_reciprocal()
}

# See what happens when you pass it malformed na.rm
calc_harmonic_mean(std_and_poor500$pe_ratio, na.rm = 1:5)
```

## Returning early

Sometimes, you don't need to run through the whole body of a function to get the answer. In that case you can return early from that function using return().

To check if x is divisible by n, you can use is_divisible_by(x, n) from assertive.

Alternatively, use the modulo operator, %%. x %% n gives the remainder when dividing x by n, so x %% n == 0 determines whether x is divisible by n. Try 1:10 %% 3 == 0 in the console.

To solve this exercise, you need to know that a leap year is every 400th year (like the year 2000) or every 4th year that isn't a century (like 1904 but not 1900 or 1905).

assertive is loaded.

```{r}
is_leap_year <- function(year) {
  # If year is div. by 400 return TRUE
  if(is_divisible_by(year, 400)) {
    return(TRUE)
  }
  # If year is div. by 100 return FALSE
  if(is_divisible_by(year, 100)) {
    return(FALSE)
  }  
  # If year is div. by 4 return TRUE
   if(is_divisible_by(year, 4)) {
    return(TRUE)
  
  
  
  # Otherwise return FALSE
   } else return(FALSE)
}
is_leap_year(year = 1900)
```

## Returning invisibly

When the main purpose of a function is to generate output, like drawing a plot or printing something in the console, you may not want a return value to be printed as well. In that case, the value should be invisibly returned.

The base R plot function returns NULL, since its main purpose is to draw a plot. This isn't helpful if you want to use it in piped code: instead it should invisibly return the plot data to be piped on to the next step.

Recall that plot() has a formula interface: instead of giving it vectors for x and y, you can specify a formula describing which columns of a data frame go on the x and y axes, and a data argument for the data frame. Note that just like lm(), the arguments are the wrong way round because the detail argument, formula, comes before the data argument.

```{r}
# Using cars, draw a scatter plot of dist vs. speed
plt_dist_vs_speed <- plot(dist ~ speed, data = cars)

# Oh no! The plot object is NULL
plt_dist_vs_speed
# Define a pipeable plot fn with data and formula args
pipeable_plot <- function(data, formula) {
  # Call plot() with the formula interface
  plot(formula, data)
  # Invisibly return the input dataset
  invisible(head(data))
}

# Draw the scatter plot of dist vs. speed again
plt_dist_vs_speed <- cars %>% 
  pipeable_plot(dist ~ speed)

# Now the plot object has a value
plt_dist_vs_speed
```

## Returning many things

Functions can only return one value. If you want to return multiple things, then you can store them all in a list.

If users want to have the list items as separate variables, they can assign each list element to its own variable using zeallot's multi-assignment operator, %\<-%.

glance(), tidy(), and augment() each take the model object as their only argument.

The Poisson regression model of Snake River visits is available as model. broom and zeallot are loaded.

```{r}
library(zeallot)
library(broom)
model <- glm(n_visits ~ gender + income + travel, 
             data =snake_river_visits )
# From previous step
groom_model <- function(model) {
  list(
    model = glance(model),
    coefficients = tidy(model),
    observations = augment(model)
  )
}

# Call groom_model on model, assigning to 3 variables

c(mdl, cff,  obs) %<-% groom_model(model)

# See these individual variables
mdl; cff; obs
```

## Returning metadata

Sometimes you want the return multiple things from a function, but you want the result to have a particular class (for example, a data frame or a numeric vector), so returning a list isn't appropriate. This is common when you have a result plus metadata about the result. (Metadata is "data about the data". For example, it could be the file a dataset was loaded from, or the username of the person who created the variable, or the number of iterations for an algorithm to converge.)

In that case, you can store the metadata in attributes. Recall the syntax for assigning attributes is as follows.

attr(object, "attribute_name") \<- attribute_value

```{r}
pipeable_plot <- function(data, formula) {
  plot(formula, data)
  # Add a "formula" attribute to data
  attr(data, "formula") <- formula
  
  invisible(data)
}

# From previous exercise
plt_dist_vs_speed <- cars %>% 
  pipeable_plot(dist ~ speed)

# Examine the structure of the result
plt_dist_vs_speed
```

## Creating and exploring environments

Environments are used to store other variables. Mostly, you can think of them as lists, but there's an important extra property that is relevant to writing functions. Every environment has a parent environment (except the empty environment, at the root of the environment tree). This determines which variables R know about at different places in your code.

Facts about the Republic of South Africa are contained in capitals, national_parks, and population.

```{r, eval=TRUE}
# From previous steps
rsa_lst <- list(
  capitals = c("J Burg", "Capetown"),
  national_parks = c("Krug", "Iddo"),
  population = 30000
)
rsa_env <- list2env(rsa_lst)

# Find the parent environment of rsa_env
parent <- parent.env(rsa_env)

# Print its name
environmentName(parent)
```

Do variables exist? If R cannot find a variable in the current environment, it will look in the parent environment, then the grandparent environment, and so on until it finds it.

rsa_env has been modified so it includes capitals and national_parks, but not population.

```{r}
# Compare the contents of the global environment and rsa_env
ls.str(globalenv())
ls.str(rsa_env)

# Does population exist in rsa_env?
exists("population", envir = rsa_env)

# Does population exist in rsa_env, ignoring inheritance?
exists("population", envir = rsa_env, inherits = FALSE)

```

## Converting areas to metric 1

In this chapter, you'll be working with grain yield data from the United States Department of Agriculture, National Agricultural Statistics Service. Unfortunately, they report all areas in acres. So, the first thing you need to do is write some utility functions to convert areas in acres to areas in hectares.

To solve this exercise, you need to know the following:

There are 4840 square yards in an acre. There are 36 inches in a yard and one inch is 0.0254 meters. There are 10000 square meters in a hectare.

```{r}
# Write a function to convert acres to sq. yards
acres_to_sq_yards <- function(acres) {
  acres * 4840
}

# Write a function to convert yards to meters
yards_to_meters <- function(yards){
    yards * 36*0.0254
}

# Write a function to convert sq. meters to hectares
# Write a function to convert yards to meters
sq_meters_to_hectares <- function(sq_meters){
    sq_meters/10000
}
```

## Converting areas to metric 2

You're almost there with creating a function to convert acres to hectares. You need another utility function to deal with getting from square yards to square meters. Then, you can bring everything together to write the overall acres-to-hectares conversion function. Finally, in the next exercise you'll be calculating area conversions in the denominator of a ratio, so you'll need a harmonic acre-to-hectare conversion function.

Free hints: magrittr's raise_to_power() will be useful here. The last step is similar to Chapter 2's Harmonic Mean.

The three utility functions from the last exercise (acres_to_sq_yards(), yards_to_meters(), and sq_meters_to_hectares()) are available, as is your get_reciprocal() from Chapter 2. magrittr is loaded.

```{r}
# Write a function to convert sq. yards to sq. meters
sq_yards_to_sq_meters <- function(sq_yards) {
  sq_yards %>%
    # Take the square root
    sqrt() %>%
    # Convert yards to meters
    yards_to_meters() %>%
    # Square it
    raise_to_power(2)
}

# Load the function from the previous step
#load_step2()

# Write a function to convert acres to hectares
acres_to_hectares <- function(acres) {
  acres %>%
    # Convert acres to sq yards
    acres_to_sq_yards() %>%
    # Convert sq yards to sq meters
    sq_yards_to_sq_meters() %>%
    # Convert sq meters to hectares
    sq_meters_to_hectares()
}

# Define a harmonic acres to hectares function
harmonic_acres_to_hectares <- function(acres) {
  acres %>% 
    # Get the reciprocal
    get_reciprocal() %>%
    # Convert acres to hectares
    acres_to_hectares %>% 
    # Get the reciprocal again
    get_reciprocal()
}
```

## Converting yields to metric

The yields in the NASS corn data are also given in US units, namely bushels per acre. You'll need to write some more utility functions to convert this unit to the metric unit of kg per hectare.

Bushels historically meant a volume of 8 gallons, but in the context of grain, they are now defined as masses. This mass differs for each grain! To solve this exercise, you need to know these facts.

One pound (lb) is 0.45359237 kilograms (kg). One bushel is 48 lbs of barley, 56 lbs of corn, or 60 lbs of wheat. magrittr is loaded.

```{r}
library(magrittr)
# Write a function to convert lb to kg
lbs_to_kgs <- function(lbs){
    lbs * 0.45359237
}

# Write a function to convert bushels to lbs
bushels_to_lbs <- function(bushels, crop) {
  # Define a lookup table of scale factors
  c(barley = 48, corn = 56, wheat = 60) %>%
    # Extract the value for the crop
    magrittr::extract(crop) %>%
    # Multiply by the no. of bushels
    multiply_by(bushels)
}


# Write a function to convert bushels to kg
bushels_to_kgs <- function(bushels, crop) {
  bushels %>%
    # Convert bushels to lbs for this crop
    bushels_to_lbs(crop) %>%
    # Convert lbs to kgs
    lbs_to_kgs()
}



# Write a function to convert bushels/acre to kg/ha
bushels_per_acre_to_kgs_per_hectare <- function(bushels_per_acre, crop = c("barley", "corn", "wheat")) {
  # Match the crop argument
  crop <- match.arg(crop)
  bushels_per_acre %>%
    # Convert bushels to kgs for this crop
    bushels_to_kgs(crop) %>%
    # Convert harmonic acres to ha
    harmonic_acres_to_hectares()
}
```

## Applying the unit conversion

Now that you've written some functions, it's time to apply them! The NASS corn dataset is available, and you can fortify it (jargon for "adding new columns") with metrics areas and yields.

This fortification process can also be turned in to a function, so you'll define a function for this, and test it on the NASS wheat dataset.

```{r}

corn <- readRDS("nass.corn.rds")
wheat <- readRDS("nass.wheat.rds")
glimpse(corn)

corn %>%
  # Add some columns
  mutate(
    # Convert farmed area from acres to ha
    farmed_area_ha = acres_to_hectares(farmed_area_acres),
    # Convert yield from bushels/acre to kg/ha
    yield_kg_per_ha = bushels_per_acre_to_kgs_per_hectare(
      yield_bushels_per_acre,
      crop = "corn"
    )
  ) %>%
  head()

# Wrap this code into a function
fortify_with_metric_units <- function(data, crop){


  data %>%
    mutate(
      farmed_area_ha = acres_to_hectares(farmed_area_acres),
      yield_kg_per_ha = bushels_per_acre_to_kgs_per_hectare(
        yield_bushels_per_acre, 
        crop = crop
      )
    )
}

# Try it on the wheat dataset
wheat <- fortify_with_metric_units(wheat, crop = "wheat")
```

## Plotting yields over time

Now that the units have been dealt with, it's time to explore the datasets. An obvious question to ask about each crop is, "how do the yields change over time in each US state?" Let's draw a line plot to find out! ggplot2 is loaded, and corn and wheat datasets are available with metric units.

```{r}
# Wrap this plotting code into a function
plot_yield_vs_year <- function(data){
  ggplot(data, aes(year, yield_kg_per_ha)) +
    geom_line(aes(group = state)) +
    geom_smooth()
}

# Test it on the wheat dataset
plot_yield_vs_year(wheat)
```

## A nation divided

The USA has a varied climate, so we might expect yields to differ between states. Rather than trying to reason about 50 states separately, we can use the USA Census Regions to get 9 groups.

The "Corn Belt", where most US corn is grown is in the "West North Central" and "East North Central" regions. The "Wheat Belt" is in the "West South Central" region.

dplyr is loaded, the corn and wheat datasets are available, as is usa_census_regions.

```{r, eval=T}
# Inner join the corn dataset to usa_census_regions by state
library(data.table)
usa_census_regions <- read_csv("usa_census_regions.csv")
setnames(usa_census_regions, c("State", "Region"), c("state", "census_region"))
View(usa_census_regions)
# corn %>%
#   inner_join(usa_census_regions, by= "state")
# Wrap this code into a function
fortify_with_census_region <- function(data){
  data %>%
    inner_join(usa_census_regions, by = "state")
}

# Try it on the wheat dataset
wheat <- fortify_with_census_region(wheat)
```

## Plotting yields over time by region

So far, you have a function to plot yields over time for each crop, and you've added a census_region column to the crop datasets. Now you are ready to look at how the yields change over time in each region of the USA.

ggplot2 is loaded. corn and wheat have been fortified with census regions. plot_yield_vs_year() is available.

```{r, eval=T}

# Wrap this code into a function
plot_yield_vs_year_by_region <- function(data) {

  plot_yield_vs_year(data) +
    facet_wrap(vars(census_region))
}

# Try it on the wheat dataset

plot_yield_vs_year_by_region(wheat)

```

## Running a model

The smooth trend line you saw in the plots of yield over time use a generalized additive model (GAM) to determine where the line should lie. This sort of model is ideal for fitting nonlinear curves. So we can make predictions about future yields, let's explicitly run the model. The syntax for running this GAM takes the following form.

gam(response \~ s(explanatory_var1) + explanatory_var2, data = dataset) Here, s() means "make the variable smooth", where smooth very roughly means nonlinear.

mgcv and dplyr are loaded; the corn and wheat datasets are available.

```{r}
# Wrap the model code into a function
library(mgcv)
run_gam_yield_vs_year_by_region <- function(data){


  gam(yield_kg_per_ha ~ s(year) + census_region, data = data)

}

# Try it on the wheat dataset
wheat_model <- run_gam_yield_vs_year_by_region(wheat)
wheat_model
```

## Making yield predictions

The fun part of modeling is using the models to make predictions. You can do this using a call to predict(), in the following form.

predict(model, cases_to_predict, type = "response") mgcv and dplyr are loaded; GAMs of the corn and wheat datasets are available as corn_model and wheat_model. A character vector of census regions is stored as census_regions.

```{r}
census_regions <- wheat$census_region %>% unique()
# Wrap this prediction code into a function
predict_yields <- function(model, year){

  predict_this <- data.table(
    year = year,
    census_region = census_regions
  ) 
  pred_yield_kg_per_ha <- predict(model, predict_this, type = "response")
  predict_this %>%
    mutate(pred_yield_kg_per_ha = pred_yield_kg_per_ha)
}

# Try it on the wheat dataset
predict_yields(wheat_model, year = 2050)
```

## Do it all over again

Hopefully, by now, you've realized that the real benefit to writing functions is that you can reuse your code easily. Now you are going to rerun the whole analysis from this chapter on a new crop, barley. Since all the infrastructure is in place, that's less effort than it sounds!

Barley prefers a cooler climate compared to corn and wheat and is commonly grown in the US mountain states of Idaho and Montana.

dplyr and ggplot2, and mgcv are loaded; fortify_with_metric_units(), fortify_with_census_region(), plot_yield_vs_year_by_region(), run_gam_yield_vs_year_by_region(), and predict_yields() are available.

```{r, eval=F}
fortified_barley <- barley %>% 
  # Fortify with metric units
  fortify_with_metric_units() %>%
  # Fortify with census regions
  fortify_with_census_region()

# See the result
glimpse(fortified_barley)
# From previous step
fortified_barley <- barley %>% 
  fortify_with_metric_units() %>%
  fortify_with_census_region()

fortified_barley %>% 
  # Run a GAM of yield vs. year by region
  run_gam_yield_vs_year_by_region %>% 
  # Make predictions of yields in 2050
  predict_yields(year = 2050)
```
